#Load the activity.csv file

```{r}
library(tidyverse)
library(knitr)
library(dplyr)
library(ggplot2)
act <- read.csv("activity.csv", header = TRUE, sep = ",")
summary(act)
```


#calculate the total no of steps taken per day and plotting them by using histogram

```{r}
act_byDay <- act %>% 
                       group_by(date)%>%
                       summarise(total_steps = sum(steps))
act_byDay$date <- as.Date(act_byDay$date)
ggplot(act_byDay, aes(x = date, y = total_steps)) +
        geom_histogram(stat = "identity")
```

#claculate the mean and median off total no of steps taken per day

```{r}
mean(act_byDay$total_steps, na.rm = TRUE)

median(act_byDay$total_steps, na.rm = TRUE)

act_mean <- act %>%
        group_by(interval) %>%
        summarise(mean_steps = mean(steps, na.rm = TRUE))
```


#find the average steps over time series plot



```{r}
ggplot(act_mean, aes(x = interval, y = mean_steps)) +
        geom_line()
```

#calculate the number of rows containing in activity data

```{r}
sum(is.na(act))
```


#impute the NA vlaues by using mean_steps fill in the NA-containing rows

```{r}
for(i in 1:nrow(act)){
        if(is.na(act$steps[i])){
                act$steps[i] = act_mean$mean_steps[which(act_mean$interval == act$interval[i])]
        }
}
sum(is.na(act$steps))
```


#after imputing ploting the histogram of total no of steps per day

```{r}
act$date = as.Date(act$date)
ggplot(act, aes(x = date, y = steps))+
        geom_histogram(stat = "identity")
```


#find mean and median of the total number of steps per day(after imputing)

```{r}
newAct_byDay <- act %>% 
                       group_by(date)%>%
                       summarise(total_steps = sum(steps))

mean(newAct_byDay$total_steps)
median(newAct_byDay$total_steps)
```


#create new variable containing weekday and weekend

```{r}
act <- act %>% mutate(typeofDay = ifelse(weekdays(act$date) == "Saturday" | weekdays(act$date) == "Sunday", "weekend", "weekday"))
table(act$typeofDay)
newAct_mean <- act %>%
        group_by(interval, typeofDay) %>%
        summarise(mean_steps = mean(steps, na.rm = TRUE))
```


#plot the avg no of steps taken per each day using panels weekday and weekend

```{r}
ggplot(newAct_mean, aes(x = interval, y = mean_steps)) +
        geom_line() +
        facet_grid(typeofDay~.)
```


